Append-only Datastore

### 2020-06-01 paper/paper.tex
Finished the rough draft.

We also gain the advantages of some more traditionally eventually consistent
systems, such as great partition tolerance \cite{dynamo}. Together, these
properties are highly desirable for many append-only applications. Consider the
following:

@@

\begin{itemize}
	\item Handing \texttt{append} and \texttt{get} requests from clients.
	\item Ordering appends.
	\item Updating the master about appended data.
	\item Receiving the eventual, global ordering of data from the master
		and imposing it on data.
	\item Syncing with other followers to communicate updates and orderings.
\end{itemize}

@@

\texttt{append} requests are also fast, limited by the speed of a local disk
write. When a client appends data to key \texttt{k}, the follower writes that
data - along with its index - to local disk before returning to the client. This
ensures that data is not lost upon crashing and rebooting. For increased fault
tolerance, the append should be written to multiple disks locally or to a remote
disk. Note that throughout this section when we refer to updating local storage,
the in-memory store is also updated immediately after data is written to disk.

@@

Followers use the \texttt{sync} request to get data they are missing,
effectively ``patching the holes" in their local datastore. This ensures the
eventually consistent property of the system.

@@

\subsection{Leader}
% Explain what happens when this is partitioned.
% Reference GFS and not transporting the data itself.
It synchronizes stuff.

\subsection{Clients}
% Describe the API and guarantees exposed to clients.
We have very distinguished clientele.

\subsection{Corner Cases}
% Multiple simultaneous writes from c1 and c2 to f1. How do we ensure that the
% GL does not order w2 before w1?
% What if the sync request can be partially satisfied. Do we not respond, or
% respond partially and expect another sync request from the same follower?

\section{Implementation}

\section{Evaluation}
We tested using some super cool gcloud stuff.

\subsection{Workload 1}
Oh wow, we're so fast.

\subsection{Workload 2}
Oh no, we're unexpectedly not that fast.

\section{Related Work}
% Talk about Kafka
This is the most filler of all filler sections. Look, a reference! \cite{dynamo}

\section{Future Work}
% Use a raft group instead of a single node for each follower and GL.
% Client fallback to other followers.
Make it actually work.

\section{Conclusions}
I'd just like everyone to pat themselves on the back here.

>>>>

We also gain the advantages of some more traditionally eventually consistent
systems, such as great partition tolerance \cite{Dynamo}. Together, these
properties are highly desirable for many append-only applications. Consider the
following:

@@

\begin{itemize}
	\item Handing \texttt{append} and \texttt{get} requests from clients.
	\item Ordering appends.
	\item Updating the leader about appended data.
	\item Receiving the eventual, global ordering of data from the leader
		and imposing it on data.
	\item Syncing with other followers to communicate updates and orderings.
\end{itemize}

@@

\texttt{append} requests are also fast, limited by the speed of a local disk
write. When a client appends data to key \texttt{k}, the follower writes that
data - along with its index - to local disk before returning to the client. This
ensures that data is not lost upon crashing and rebooting. For increased fault
tolerance, the append should be written to multiple disks locally or to a remote
disk as well. Note that throughout this section when we refer to updating local
storage, the in-memory store is also updated immediately after data is written
to disk.

@@

Followers use the \texttt{sync} request to get data they are missing,
effectively ``patching the holes" in their local datastore. This ensures the
eventually consistent property of the system. \texttt{sync} requests also signal
what data the sender has to the recipient, enabling the recipeint to also patch
holes in its local datastore.

@@

\subsection{Leader}
% TODO: Explain what happens when this is partitioned. Look at gfs paper.
The leader is responsible for globally ordering writes for each key. Like GFS's
master \cite{GFS}, it only stores metadata rather than the datastore itself.
This decreases network utilization, as data itself is never sent.  Also, the
burden on disks is lighter per update, increasing writes per second and the
average lifetime of each disk. Like GFS, the simplicity of a single master
simplifies our implementation and the rest of our design, both of which increase
the overall stability of the system.

% TODO: Add the nonce.
The leader answers \texttt{update} requests from followers. These requests
contain the key being appended to, the index at which the follower is trying
to append, and a nonce to prevent erroneous double-appends. The leader
authoritatively orders appends by keeping a map of key to current index. If a
mapping exists from key $k$ to index $i$, an update specifying a write to index
$i+1$ increments $i$ and returns \texttt{SUCCESS}. The leader writes all map
updates to disk before responding.

An \texttt{update} to $i' < i$ indicates that the sender is not synced to index
$i$. In this case the leader increments its index and reassigns the append the
index of $i+1$. It communicates this back to the follower, and tells it to
sync with the follower that originated the append at index $i$.

Along with the current index, the leader's map also contains the follower that
originated the append at that index. This allows followers to be directed at the
freshest information whether

\subsubsection{Broadcasting Updates}
Although the system as described is eventually consistent, it may lead to
unacceptably stale data. An updated value might get synced between followers
$a$ and $b$, but follower $c$ would miss the update entirely. If a value is
updated and 3 days later follower $c$ syncs via a \texttt{get} or
\texttt{append}, $c$'s client will get 3-day-old data.

To prevent excessive data staleness, the leader is able to broadcast updates to
every follower. Depending on the workload, broadcasting every update may slow
down the system and excessively consume bandwidth. Thus users can pass a flag
specifying whether they want broadcasts for every update or periodic broadcasts.
New broadcast behavior is easy to add, as it requires implementing a
single-method Go interface that can make use of the same utility functions we
used to write our broadcasters.

\subsubsection{Leader Fault Tolerance}
Because the leader writes all updates to disk, it tolerates crashes and reboots.
For greater fault tolerance, it should write to multiple disks or a remote disk
as well.

The system continues to function when the leader is partitioned, but followers
and thus clients do not receive updates from other followers. As long as the
partition is eventually healed, the system will propgate information correctly
and self-heal. If for some reason there is a permanent partition, it must be
manually worked around by changing the cluster configuration (i.e. the set of
nodes in the system).

\subsection{Clients}
The append-only datastore presents clients with a simple interface consisting of
the \texttt{get} and \texttt{append} API described above. The consistency model
guarantees that clients:

\begin{itemize}
	\item Can read their own appends immiedately.
	\item Never see data reordered relatively.
	\item Communicate only with a fast, nearby follower, and no other nodes
		in the system.
	\item Can always read and write, even when their follower is partitioned
		from the rest of the system.
\end{itemize}

The sacrifices made are:

\begin{itemize}
	\item Consistency is eventual, not stronger.
	\item We do not support random writes.
	\item The system continues to work but may provide stale data in the
		event of a leader failure.
\end{itemize}

\section{Evaluation}
The system is written entirely in Go and can be run manually or distributed and
run with Docker. Nodes communicate via GRPC.

We ran tests with a cluster of Google Compute Engine virtual
machines. The cluster consisted of:

\begin{itemize}
	\item A leader in Seoul, Korea.
	\item A follower and client in Frankfurt, Germany.
	\item A follower and client in Los Angeles, USA.
	\item A follower and client in São Paulo, Brazil.
\end{itemize}

\subsection{Workload 1}
We should probably read and write a lot.

\subsection{Workload 2}
Let's see what happens if we kill a follower.

\section{Related Work}
Amazon's Dynamo \cite{Dynamo} supports always-writable semantics, high partition
tolerance, and laser-focuses on low-latency operation. Unlike our system, Dynamo
is a key-value store. It also exposes a great deal of complexity to developers,
who have to tailor their use of Dynamo such that conflicts are resolvable and
must manually implement conflict resolution in clients. Our system does not
allow for conflicts.

Google's GFS \cite{GFS} is also optimized for append-heavy workloads and uses a
single master for simplicity and intelligent coordination. It supports random
writes as well. However, it is explicitly optimized for non-latency-sensitive
applications, and a single read can require multiple hops (the GFS master and
chunkserver).

% TODO: Make this better.
LinkedIn's Kafka \cite{Kafka} also provides eventual delivery of large
quantities of data (specifically logs), but contains a publish/subscription
mechanism.

\section{Future Work}
We could greatly increase throughput by providing a client library (rather than
a raw GRPC interface) that is aware of the version or latest verions it holds.
This would enable followers to selectively return only the few missing pieces of
data a client is missing rather than all the data.

To increase fault tolerance, we have discussed making each follower and the
leader their own small cluster of concensus nodes. This would greatly increase
fault tolerance and, while it may complicate the system, would not complicate
clients or the protocols via which nodes communicate.

Lastly, there may be cases where followers fail or are partitioned from their
nearby clients. In these cases, we would like to explore whether clients can
fall back to another follower.

\section{Conclusions}
We built a datastore to provide low latency and high availability for append-only
data storage. While this work is tailed to a specific set of workloads, there
are numerous applications that can benefit from this approach.

Our system can be used to support distributed services such as distributed
monitoring and chat applications with high performance. Importantly, it exposes
a simple interface allowing for simple clients. We believe that this makes it a
useful tool in building distributed and microservice systems.
